---
title: "Capstone - Milestone Report"
author: "Jill Beck"
date: "April 23, 2016"
output: html_document
---

## Goal and Motivation

The goal of this project is just to display that you've gotten used to working with the data and that you are on track to create your prediction algorithm. Please submit a report on R Pubs (http://rpubs.com/) that explains your exploratory analysis and your goals for the eventual app and algorithm. This document should be concise and explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. You should make use of tables and plots to illustrate important summaries of the data set. 

The motivation for this project is to:

1. Demonstrate that you've downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app.

The data provided for NLP (Natural Language Processing) consists of 3 "corpora" of data:

* Blog posts
* News articles
* "Tweets" on Twitter

## Load Requisite Libraries, Download/Extract Data, Read Data

```{r}
# Load libraries and suppress messages for ease of reading report
suppressMessages(library(dplyr)) 
suppressMessages(library(ggplot2)) 
suppressMessages(library(LaF)) 
suppressMessages(library(quanteda)) 
suppressMessages(library(RColorBrewer)) 
suppressMessages(library(RWeka)) 
suppressMessages(library(SnowballC))
suppressMessages(library(tau)) 
suppressMessages(library(tm)) 
suppressMessages(library(wordcloud))
```

## Data: Download, Extract and Read
```{r}
# Download and extract data
source_file <- "http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
destination_file <- "Coursera-SwiftKey.zip"
download.file(source_file, destination_file)
unzip(destination_file)

# Unzip file
unzip(destination_file, list = TRUE )
```

```{r}
# Load the data en_US data
dataBlogs <- readLines("./final/en_US/en_US.blogs.txt", encoding = "UTF-8", skipNul = TRUE)
dataNews <- readLines("./final/en_US/en_US.news.txt", encoding = "UTF-8", skipNul = TRUE, warn = FALSE)
dataTwitter <- readLines("./final/en_US/en_US.twitter.txt", encoding = "UTF-8", skipNul = TRUE)

#Convert to ASCII
dataNews <- iconv(dataNews, 'UTF-8', 'ASCII', "byte")
dataBlogs <- iconv(dataBlogs, 'UTF-8', 'ASCII', "byte")
dataTwitter <- iconv(dataTwitter, 'UTF-8', 'ASCII', "byte")
```

## Exploratory Analysis, inc. Word Counts, Line Counts

```{r}
# Assess size of all 3 files - blogs, news and Twitter
dataBlogs.filesizeMB <- file.size("./final/en_US/en_US.blogs.txt")
dataNews.filesizeMB <- file.size("./final/en_US/en_US.news.txt")
dataTwitter.filesizeMB <- file.size("./final/en_US/en_US.twitter.txt")
```

```{r}
# Determine word count and length of the longest line seen
dataBlogs.wordsCount <- nchar(dataBlogs)
tmax <- which.max(dataBlogs.wordsCount)
dataBlogs.longestWordCount <- nchar(dataBlogs[tmax])

dataNews.wordsCount <- nchar(dataNews)
tmax <- which.max(dataNews.wordsCount)
dataNews.longestWordCount <- nchar(dataNews[tmax])

dataTwitter.wordsCount <- nchar(dataTwitter)
tmax <- which.max(dataTwitter.wordsCount)
dataTwitter.longestWordCount <- nchar(dataTwitter[tmax])
```

```{r}
# Combine into a data frame
dataframe.blogs <- c(dataBlogs.filesizeMB, length(dataBlogs.wordsCount), dataBlogs.longestWordCount)
dataframe.news <- c(dataNews.filesizeMB, length(dataNews.wordsCount), dataNews.longestWordCount)
dataframe.twitter <- c(dataTwitter.filesizeMB, length(dataTwitter.wordsCount), dataTwitter.longestWordCount)
```

### Create Table: File Size, Word Count and Longest Line

```{r}
info <- data.frame(rbind(dataframe.blogs, dataframe.news, dataframe.twitter))
names(info) <- c("File Size (MB)", "Word Count", "Longest Line")
row.names(info) <- c("Blogs", "News", "Twitter")

# Showcase table
info
```

### Sampling Exercise: Word Frequency

```{r}
# Assess maximum number of characters in a line of the files
summary(nchar(dataBlogs))[6] 
summary(nchar(dataNews))[6] 
summary(nchar(dataTwitter))[6] 

# Run sampling at 5% of the actual file parameters because of sizes of files
dataBlogs_sample_size   <- round(.05 * length(dataBlogs), 0)
dataNews_sample_size    <- round(.05 * length(dataNews), 0) 
dataTwitter_sample_size <- round(.05 * length(dataTwitter), 0)

# Compute with approximately 5% of the population for each file
dataBlogs_sample <- sample_lines("./final/en_US/en_US.blogs.txt", n = dataBlogs_sample_size, nlines = NULL) 
dataNews_sample <- sample_lines("./final/en_US/en_US.news.txt", n = dataNews_sample_size , nlines = NULL) 
dataTwitter_sample <- sample_lines("./final/en_US/en_US.twitter.txt", n = dataTwitter_sample_size, nlines = NULL)

# Determine word frequency for each of the 3 files
dataBlogs_word_freq <- dfm(dataBlogs_sample, verbose = FALSE)
dataNews_word_freq <- dfm(dataNews_sample, verbose = FALSE)
dataTwitter_word_freq <- dfm(dataTwitter_sample, verbose = FALSE)

docfreq(dataBlogs_word_freq)[1:11]
docfreq(dataNews_word_freq)[1:11]
docfreq(dataTwitter_word_freq)[1:11]
```

## Clean Corpus & Incorporate "Bad Words"

```{r}
require(tm)
require(SnowballC)
require(RWeka)
require(slam)
require(ggplot2)
```

```{r}
# Set CleanR function
CleanR <- function(corpus){
        tm_map(corpus, removeNumbers) %>%
                tm_map(removePunctuation) %>%
                tm_map(content_transformer(tolower)) %>%
                tm_map(stripWhitespace) %>%
                tm_map(stemDocument)
                tm_map(PlainTextDocument)
}
```

```{r}
# Combine all twitter, blogs and news and save to RData
all <- c(dataBlogs, dataNews, dataTwitter)
save(all, file="all.RData")

all.sample <- sample(all, round(0.02*length(all)))
save(all.sample, file="sample-2p.RData")
```

```{r}
# Create the corpus
corpus <- Corpus(VectorSource(all.sample))
corpus <- tm_map(corpus, content_transformer(removePunctuation), lazy = TRUE)
corpus <- tm_map(corpus, content_transformer(removeNumbers), lazy = TRUE)
corpus <- tm_map(corpus, content_transformer(tolower), lazy = TRUE)
corpus <- tm_map(corpus, content_transformer(stripWhitespace), lazy = TRUE)
corpus <- tm_map(corpus, content_transformer(PlainTextDocument), lazy = TRUE)
```

```{r}
# Set CleanR function
CleanR <- function(corpus){
        tm_map(corpus, removeNumbers) %>%
                tm_map(removePunctuation) %>%
                tm_map(content_transformer(tolower)) %>%
                tm_map(stripWhitespace) %>%
                tm_map(stemDocument)
}
```

```{r}

# reading the text file
bad.word <- read.delim(file = "googlebadwords.txt", sep = ":", header = FALSE)
bad.word_new<- gsub("[*()]","",bad.word[,1])
corpus <- tm_map(corpus, removeWords, bad.word_new)
```

```{r}
# Save the corpus for next phase of capstone
save(corpus, file="WorkingCorpus.RData")
```

## NGram Analysis

```{r}
# Create a few NGram functions via RWeka
unigram_token <- function(x)  NGramTokenizer(x, Weka_control(min = 1, max = 1))
bigram_token <- function(x)   NGramTokenizer(x, Weka_control(min = 2, max = 2))
trigram_token <- function(x)  NGramTokenizer(x, Weka_control(min = 3, max = 3))
```

### UniGram analysis, inc. Histogram

```{r}
# Create UniGram functions via RWeka
options(stringsAsFactors = FALSE)
options(mc.cores = 1)
unigram <- TermDocumentMatrix(corpus, control=list(tokenize=unigram_token))
unigram.good <- rollup(unigram, 2, na.rm=TRUE, FUN = sum)
```

```{r}
# Sort with decreasing frequency
unigram.tf <- findFreqTerms(unigram.good, lowfreq = 3)
unigram.tf <- sort(rowSums(as.matrix(unigram.good[unigram.tf, ])), decreasing = TRUE)
unigram.tf <- data.frame(unigram.good=names(unigram.tf), frequency=unigram.tf)
names(unigram.tf) <- c("word", "frequency")
head(unigram.tf, 10)
```

```{r}
# Plot top 10 word frequency for UniGram
g <- ggplot(data = head(unigram.tf, 10), aes(x = word, y = frequency))
g <- g + geom_bar(stat="Identity", fill="red", colour = "black")
g <- g + geom_text(aes(label=frequency), vjust=-0.1)
g <- g + theme(axis.text.x = element_text(angle = 45, hjust = 2))
g
```

### BiGram Analysis
```{r}
# BiGram work
bi.gram.dataBlogs <- textcnt(dataBlogs_sample, n = 2, method = "string") 
bi.gram.dataBlogs <- bi.gram.dataBlogs[order(bi.gram.dataBlogs, decreasing = TRUE)]
bi.gram.dataBlogs[1:3] # top three, 2-Word combinations
```

## Word Clouds

```{r}
blogs_corpus <- VCorpus(DataframeSource(data.frame(dataBlogs_sample)))
news_corpus <- VCorpus(DataframeSource(data.frame(dataNews_sample)))
twitter_corpus <- VCorpus(DataframeSource(data.frame(dataTwitter_sample)))

rm(dataBlogs_sample); rm(dataNews_sample); rm(dataTwitter_sample)

blogs_corpus <- CleanR(blogs_corpus)
news_corpus <- CleanR(news_corpus)
twitter_corpus <- CleanR(twitter_corpus)

pal <- brewer.pal(8,"Accent")

wordcloud(blogs_corpus, max.words = 90, random.order = FALSE, colors = pal)
wordcloud(news_corpus, max.words = 90, random.order = FALSE, colors = pal)
wordcloud(twitter_corpus, max.words = 90, random.order = FALSE, colors = pal)
```

## Next Steps

#### Create a Shiny app that takes as input a phrase (multiple words) in a text box input and outputs a prediction of the next word. 

#### Develop report to showcase findings and document how Shiny app works.